% FILENAME: paper/section-Matter.tex

\todo{Author Overview: This section proves that particles are not fundamental points added to the spacetime stage, but rather topological defects (knots) within the connection itself.
Order of theorems covered:
1. kinematicMultiColorRequirement / kinematicSingleColorDegeneracy: Abelian fields collapse space. Volume requires non-Abelian fields (Color).
2. kinematicTopologicalStability: Instantons cannot decay due to Cartan-Maurer bounds.
3. topologicalMassGap: SU(2) curvature strictly generates positive inertial mass.
4. kinematicStringConfinement: 3D geometric strings crush into the exact 1D confinement Hamiltonian.
5. macroscopicVolumeImpliesAxialCondensate: Broken chiral symmetry forces an isovector, parity-odd axial condensate.
Nuances: Address the "SU(3) Problem"—why the Strong Force emerges geometrically from SU(2) topology without needing a fundamental SU(3) gauge group.}

\todo{Author: Review and rename section title if desired.}
\section{Matter: Topological Knots and the Mass Gap}
\label{sec:Matter}

\todo{Author: Review and rename subsection title if desired.}
\subsection{The Necessity of Non-Abelian Fields}
\label{subsec:non_abelian}

\todo{Author: Explain why a U(1) Abelian field (like pure electromagnetism) cannot create space on its own. The volume determinant natively collapses. The universe absolutely requires non-Abelian fields (color) to expand into a macroscopic geometry.}

\begin{equation}
    [\mathcal{F}_{\mu\nu}, \mathcal{F}_{\rho\sigma}] = 0 \implies \det(g_{\mu\nu}) = 0
    \quad \leanref{CGD.Particles.Color.kinematicSingleColorDegeneracy}
    \label{eq:kinematicSingleColorDegeneracy}
\end{equation}

\begin{equation}
    \det(g_{\mu\nu}) \neq 0 \implies \text{Field is strictly Non-Abelian}
    \quad \leanref{CGD.Particles.Color.kinematicMultiColorRequirement}
    \label{eq:kinematicMultiColorRequirement}
\end{equation}

\todo{Author: Review and rename subsection title if desired.}
\subsection{Topological Stability and the Mass Gap}
\label{subsec:topological_stability}

\todo{Author: Explain that particles are stable topological knots (instantons) in the gauge field. Introduce the Cartan-Maurer degree / Belavin bounds to explain why these knots cannot mathematically decay or unwrap into the trivial vacuum.}

\begin{equation}
    \int_{S^3} \text{Cartan-Maurer Form} \neq 0 \implies \mathcal{A} \text{ cannot continuously decay to } 0
    \quad \leanref{CGD.Particles.TopologicalStability.kinematicTopologicalStability}
    \label{eq:kinematicTopologicalStability}
\end{equation}
\todo{Cite Litlib Dependency: \cite{belavin1975pseudoparticle} for the topological energy bounds and \cite{nakahara2003geometry} for the Cartan-Maurer degree.}

\todo{Author: Explain how the trace of the SU(2) curvature defect mathematically forces the inertial mass of these knots to be strictly positive, proving the Mass Gap from geometry.}

\begin{equation}
    \text{Tr}(\mathcal{F}^{(R)}_{\mu\nu} \mathcal{F}^{(R)\mu\nu}) \neq 0 \implies m_{\text{inertial}} > 0
    \quad \leanref{CGD.Particles.Mass.topologicalMassGap}
    \label{eq:topologicalMassGap}
\end{equation}

\todo{Author: Review and rename subsection title if desired.}
\subsection{Geometric Confinement and Phenomenology}
\label{subsec:geometric_confinement}

\todo{Author: Explain that multi-color components natively crush into 1D topological flux tubes (strings), perfectly yielding a 1D confining Hamiltonian.}

\begin{equation}
    \mathcal{H}_{3D} \rightarrow \frac{1}{2} E_z^2 + \frac{1}{2} \text{Tr}(B^2) \quad \text{(1D Confinement Reduction)}
    \quad \leanref{CGD.Particles.Confinement.kinematicStringConfinement}
    \label{eq:kinematicStringConfinement}
\end{equation}

\todo{Author: Note that breaking the L=R chiral symmetry to establish macroscopic volume mathematically guarantees the existence of a pervasive Axial-Vector condensate in the background.}

\begin{equation}
    \mathcal{A}^{(L)} \neq \mathcal{A}^{(R)} \implies \mathcal{A}_{\text{axial}} \neq 0
    \quad \leanref{CGD.Phenomenology.AxialCondensate.macroscopicVolumeImpliesAxialCondensate}
    \label{eq:macroscopicVolumeImpliesAxialCondensate}
\end{equation}
