% FILENAME: paper/section-Cosmology.tex

\todo{Author Overview: This section establishes that space and time are not a fixed background but a dynamic condensate. We must prove this before discussing gravity.
Order of theorems covered:
1. kinematicVacuumCondensate: Empty space collapses to zero volume.
2. kinematicStaticUniverseDegeneracy: A static universe collapses to zero volume.
3. kinematicTimeEmergence: A purely symmetric SO(4) state forbids a Lorentzian metric.
4. kinematicBigBang: The boundary state is a non-degenerate SO(4) instanton bounce.
5. kinematicParityInversion: The arrow of time geometrically maps to matter/antimatter asymmetry.
6. kinematicClassicalScaleBreaking: The emergent metric classically breaks conformal symmetry.
Nuances: Emphasize that because we lack classical equations of motion, these topological boundaries are what literally force the universe to expand and evolve.}

\todo{Author: Review and rename section title if desired.}
\section{Cosmology: The Genesis of Spacetime}
\label{sec:Cosmology}

\todo{Author: Review and rename subsection title if desired.}
\subsection{The Necessity of Dynamic Spacetime}
\label{subsec:dynamic_spacetime}

\todo{Author: In standard General Relativity, the metric is assumed to exist independently of the fields. Explain that in CGD, spacetime volume is a dynamically generated condensate. If the connection is an empty vacuum, or if it is completely static, the macroscopic volume mathematically collapses to zero. Therefore, space strictly requires dynamic time-evolution to exist.}

\begin{equation}
    \text{If } \det(g_{\mu\nu}) \neq 0 \text{ for } x \in \text{Bulk}, \quad \text{then } \mathcal{A}_\mu(x) \neq 0
    \quad \leanref{CGD.Quantum.Vacuum.kinematicVacuumCondensate}
    \label{eq:kinematicVacuumCondensate}
\end{equation}

\begin{equation}
    \text{If } \partial_0 \mathcal{A}_i = 0, \quad \text{then } \det(g_{\mu\nu}) = 0
    \quad \leanref{CGD.Cosmology.BigBang.kinematicStaticUniverseDegeneracy}
    \label{eq:kinematicStaticUniverseDegeneracy}
\end{equation}

\todo{Author: Review and rename subsection title if desired.}
\subsection{Time Emergence and the SO(4) Instanton Bounce}
\label{subsec:time_emergence}

\todo{Author: Explain that a Lorentzian time dimension emerges geometrically only by spontaneously breaking symmetric Euclidean SO(4) topology. Frame the initial boundary condition as a non-singular Euclidean instanton bounce rather than a mathematical singularity.}

\begin{equation}
    \mathcal{F}_{\mu\nu} = \frac{1}{2} \epsilon_{\mu\nu}^{\phantom{\mu\nu}\rho\sigma} \mathcal{F}_{\rho\sigma} \implies \text{Metric signature is degenerate or Euclidean}
    \quad \leanref{CGD.Cosmology.TimeEmergence.Theorem.kinematicTimeEmergence}
    \label{eq:kinematicTimeEmergence}
\end{equation}
\todo{Cite Litlib Dependency: \cite{hartle1983wave} for the phenomenological context of the no-boundary SO(4) proposal.}

\begin{equation}
    g_{\mu\nu}\Big|_{\text{Boundary}} = c \delta_{\mu\nu}, \quad c \neq 0
    \quad \leanref{CGD.Cosmology.BigBang.kinematicBigBang}
    \label{eq:kinematicBigBang}
\end{equation}

\todo{Author: Review and rename subsection title if desired.}
\subsection{Parity Inversion and Scale Breaking}
\label{subsec:parity_and_scale}

\todo{Author: Explain parity inversion mapping to topological charge. Negating the electric/temporal components of the connection exactly negates the Pontryagin density. Then briefly mention that the Urbantke metric natively breaks conformal scale symmetry, locking in a macroscopic scale.}

\begin{equation}
    \mathcal{A}_0 \to -\mathcal{A}_0 \implies \mathcal{P} \to -\mathcal{P}
    \quad \leanref{CGD.Cosmology.ParityInversion.kinematicParityInversion}
    \label{eq:kinematicParityInversion}
\end{equation}
\todo{Cite Litlib Dependency: \cite{sakharov1991violation} for the phenomenological context of matter/antimatter asymmetry.}

\begin{equation}
    \mathcal{F}_{\mu\nu} \to \lambda^2 \mathcal{F}_{\mu\nu} \implies g_{\mu\nu} \to \lambda^6 g_{\mu\nu}, \quad \det(g) \to \lambda^{24} \det(g)
    \quad \leanref{CGD.Cosmology.ScaleBreaking.kinematicClassicalScaleBreaking}
    \label{eq:kinematicClassicalScaleBreaking}
\end{equation}
